Efficient stability for the Pauli group

### 2023-06-20 efficient_stability.tex
finished a pass on Section 5

To describe the tester we need to introduce interpolation coefficients, as follows. For $u\in \F_q$ and $i\in\{1,\ldots,d\}$ let 
\[ \alpha_{u,i} = \frac{\prod_{\substack{i'=1\\i'\neq i}}^{d+1} i'}{\prod_{\substack{i'=1\\i'\neq i}}^{d+1} (i'-i)}\;.\]
These are defined so that for any $f:\F_q\to\F_q$ of degree at most $d$, $f(u)=\sum_{i=1}^{d+1} \alpha_{u,i} f(u+i)$. The tester verifies this relation along a randomly chosen direction. 
\tnote{Here I chose to specialize the ``line'' to $(d+2)$ consecutive points on it, which is notationally marginally simpler than $(d+2)$ uniformly random points on it. Hopefully it's fine}

%The second test applied by the tester, the subcube commutation test, may seem superfluous, because it always accepts. However, the test is important to show that the code is robust. Note that including the test imposes a non-trivial constraint of pairwise approximate commutation on representations of $\mC_\RM$, and hence also on $\eps$-local presentations. Both tests are easily seen to be linear; in particular, testing that $q$ entries match the truth table of a polynomial of degree at most $d$ can be done by checking $(n-d-1)$ linear equalities, because the set of evaluation vectors  of polynomials of  degree at most $d<m$ at any $m$ distinct points  is a linear subspace of dimension $(d+1)$ of $\F_q^m$.

@@

\begin{figure}[!htbp]
  \centering
  \begin{gamespec}
Given access to some $g\in (\F_2^t)^n$, where $n=q^m$, identify $g$ with a function $g:(\F_2^t)^m \times \F_2^t \to F_2$. Perform one of the following tests with probability~$\tfrac{1}{2}$ each. \tnote{Needs to be filled in}
\begin{enumerate}
	\item \textbf{Low-degree test:}
		Let $u \sim \F_q^m$ be a uniformly random point and $j\sim \{1,\ldots,m\}$ chosen uniformly at random. Let $e_j=(0,\ldots,0,1,0,\ldots,0)\in \F_q^m$, where the unique $1$ is in the $j$-th position. 
	\item \textbf{Hadamard test:} 	
    \end{enumerate}
  \end{gamespec}
  \caption{A local test for $\code_{\bRM}$}
  \label{fig:bRM-tester}
\end{figure}

We do not know if $G_\bRM = \Z_2^K$, with $K=t(d+1)^m$. Instead we modify the presentation $G(h_\bRM)$ by adding pairwise commutation relations in a similar manner as Lemma~\ref{lem:com-code}. Let 
\[ G(h_\bRM) \,=\, \big\langle x_1,\ldots,x_N \;:\; \{R^\sq_k\}\,,\; \{R^\ld_k\}\,,\; \{R^\had_{k}\} \big\rangle\;,\]
where  $R^\sq_k$ ranges overal all relations of the form $x_i^2=e$ for $i\in \{1,\ldots,N\}$, $R^\ld_k$ ranges over all relations implied by the ``low-degree test'' in Figure~\ref{fig:bRM-tester} and $R^\had_{k}$ ranges over all the relations implied by the ``Hadamard test.'' For $k=(i,j)\in\{1,\ldots,N\}$ such that $i<j$ let $R^\com_k$ be the relation $[x_i,x_j]=e$.
Then we define
\begin{equation}\label{eq:z2k-eff}
 G' \,=\,\big\langle x_1,\ldots,x_N \;:\; \{R^\sq_k\}\,,\; \{R^\ld_k\}\,,\; \{R^\had_{k}\}\, , \; \{ R^\com_k\}\big\rangle\;.
\end{equation}
From Lemma~\ref{lem:com-code} it follows that $G'=\Z_2^K$. Our main result is an efficient stability result for this presentation. To state this we need to introduce distributions $\mu_S$ and $\mu_R$ on the generators and relations of $G'$. The distribution $\mu_S$ is taken to be the uniform distribution over $[N] = (\F_2^t)^m \times \F_2^t$. The distribution $\mu_R$ is obtained as follows. With probability $1/4$ each, a relation from $\{R^\sq_k\}$, $\{R^\ld_k\}$ or $\{R^\had_k\}$ is chosen uniformly at random. With probability $1/4$, a random commutation relation from $R^\com_k$ is chosen according to the following distribution. First select $j\in\{1,\ldots,m\}$ and $u_{m-j+2},\ldots,u_m \in \F_2^t$ uniformly at random. Then select $v,v' \in (\F_2^t)^m$ uniformly at random, conditioned on the last $(j-1)$ coordinates matching $u_{m-j+2},\ldots,u_m$. Finally, select $\alpha,\beta\in \F_2^t$ uniformly at random. Check commutation between $x_{u,\alpha}$ and $x_{u,\beta}$.

@@

\begin{proof}\tnote{in progress}
Let $(\mM,\tau)$ be a tracial von Neumann algebra and $\phi$ be an $\eps$-homomorphism of $\langle S:R\rangle$ on $(\mU(\mM),\tau)$. 
 Here, $S = \{s_{u,\alpha}: u\in (\F_2^t)^m, \alpha\in \F_2^t\}$ and $R$ is the set of all relations in~\eqref{eq:z2k-eff}, i.e.\ $R= \{R^\ld_k\}\cup\{R^\had_k\}\cup\{R^\com_k\}$. We first show that we may assume without loss of generality that $\phi$ sends each element of $S$ to a Hermitian involution.

\begin{claim}\label{claim:z2-stab-1}
There is an $\eps^{(1)}=O(\eps)$ and an $\eps^{(1)}$-homomorphism $\phi^{(1)}$ of $\langle S:R\rangle$ on $(\mU(\mM),\tau)$ such that $\phi^{(1)}(x)$ is a Hermitian involution for all $x\in S$, and furthermore
\[ \Es{x\sim\mu} \big\| \phi(x) - \phi^{(1)}(x) \big\|_\tau^2 \,\leq\, \eps^{(1)}\;.\] 
\end{claim}

\begin{proof}
Using elementary calculations (see e.g.~\cite[Lemma 3.6]{slofstra2019set}) we see that for any complex $\alpha$, 
\[ \big| \sgn\Re\alpha-\alpha\big| \,\leq\, \Big(1+\frac{1}{\sqrt{2}}\Big) \big|\alpha^2 -1 \big|\;.\]
For any $i\in \{1,\ldots,N\}$ let $\phi^{(1)}(x_i) = \sgn\Re (\phi(x_i))$. Then the claim follows since 
\[ \Es{i\in\{1,\ldots,N\}} \big\| \phi(x_i)^2-\Id \big\|_\tau^2 \,\leq\, 4\eps\;,\]
by assumption and the definition of $\mu_R$. 
\end{proof}

For ease of notation we relabel $\phi^{(1)}$ and $\eps^{(1)}$ as $\phi$ and $\eps$ respectively. We first exploit the relations $\{R^\had_k\}$ to show the following.

\begin{claim}\label{claim:z2-stab-2}
Let $q=2^t$. 
For every $u\in \F_q^m$ there is a projective measurement $\{P^u_\alpha\}_{\alpha\in \F_q}$ on $\mM$ such that 
\[ \Es{u\in \F_q^m} \Es{\alpha\in \F_2^t} \big\| \phi(s_{u,\alpha}) - \sum_{\beta\in\F_2^t} (-1)^{\alpha \cdot \beta} P^u_\beta \big\|_\tau^2 \,=\, O(\eps)\;. \]
\end{claim}

\begin{proof}
Since $\mu_R$ places weight $1/4$ on relations $\{R^\had_k\}$ we deduce that 
\begin{equation}\label{eq:stab-rm-1}
\Es{u\in \F_q^m} \Es{\alpha,\beta\in \F_2^t} \big\|\phi(s_{u,\alpha})\phi(s_{u,\beta})\phi(s_{u,\alpha+\beta})-\Id\big\|_\tau^2 \,\leq\, 4\eps\;. 
\end{equation}
Fix an $u\in \F_q^m$ and apply Lemma~\ref{lem:had-stab} for that $u$. This gives a partial isometry $w\in P\mM_\infty\Id_\mM$ and a family of commuting unitaries $\{U_\alpha\}$ on $\mN=P\mM_\infty P$ such that 
\[ \Es{\alpha \in \F_2^t} \big\| \phi(s_{u,\alpha}) - w^* U_\alpha w \big\|_\tau^2 \,=\, O(\eps_u)\;,\]
where $\eps_u\geq 0$ is such that $\Es{u\in \F_q^m} \eps_u = 4\eps$. Because $\{U_\alpha\}$ are a representation of $\F_2^t$, there is a projective measurement $\{Q^u_\alpha\}_{\alpha\in \F_2^t}$ on $\mN$ such that $U_\alpha = \sum_\beta (-1)^{\alpha\cdot \beta} Q^u_\beta$. ($\{Q^u_\alpha\}$ can be found explicitly by applying the Fourier transform.) Applying Lemma~\ref{lem:pull-back}, we deduce a projective measurement $\{P^u_\alpha\}_{\alpha\in \F_2^t}$  on $\mM$ such that (by the triangle inequality)
\[ \Es{\alpha \in \F_2^t} \big\| \phi(s_{u,\alpha}) - \sum_{\beta\in\F_2^t} (-1)^{\alpha \cdot \beta} P^u_\beta \big\|_\tau^2 \,=\, O(\eps_u)\;.\]
\end{proof}

The following claim will prove useful.

\begin{claim}\label{claim:z2-stab-2b}\tnote{This will require a specific distribution on the commutation relations}
\[ \Es{u\in \F_q^m} \Es{j\in\{1,\ldots,m\}} \Es{i\neq i' \in \{0,\ldots,d+1\}} \big\| [P^{u+ie_j}_\alpha, P^{u+i'
 e_{j}}_\beta ]\big\|_\tau^2 = O(\eps)\;. \]
\end{claim}

\begin{proof}
From Claim~\ref{claim:z2-stab-2} we have that 
\[ \Es{u\in \F_q^m} \Es{\alpha\in \F_2^t} \big\|\Es{\alpha\in \F_2^t} (-1)^{\alpha \cdot \beta} \phi(s_{u,\alpha}) -  P^u_\beta \big\|_\tau^2 \,=\, O(\eps)\;. \]
The claim therefore follows from the commutation relations. 
\end{proof}

@@

\begin{claim}\label{claim:z2-stab-3}
For every axis-parallel line $\ell\subseteq \F_q^m$ there is a projective measurement $\{Q^\ell_g\}$ on $\mM$ such that 
\[ \Es{u\in \F_q^m} \sum_f \tau\big( P^u_{f(u)} Q^\ell_g\big) \,\geq\, 1-\poly(\eps)\;. \]
\end{claim}

>>>>

To describe the tester we need to introduce interpolation coefficients, as follows. For $u,v\in \F_q$ and $i\in\{0,\ldots,d\}$ let 
\begin{equation}\label{eq:interp-coeff}
 \alpha_{u,v,i} = \frac{\prod_{\substack{i'=0\\i'\neq i}}^{d} \big(v-(u+i')\big)}{\prod_{\substack{i'=0\\i'\neq i}}^{d} \big(u+i-(u+i')\big)}\;.
\end{equation}
These are defined so that for any $f:\F_q\to\F_q$ of degree at most $d$, 
\[ f(v)\,=\, \sum_{i=0}^{d} \alpha_{u,v,i} \, f(u+i)\;.\]
The tester verifies this relation along a randomly chosen direction.

@@

\begin{figure}[!htbp]
  \centering
  \begin{gamespec}
Given access to some $g\in (\F_2^t)^n$, where $n=q^m$, identify $g$ with a function $g:(\F_2^t)^m \times \F_2^t \to \F_2$. Perform one of the following tests with probability~$\tfrac{1}{2}$ each. 
\begin{enumerate}
	\item \textbf{Low-degree test:}
		Let $u \sim \F_q^m$ be a uniformly random point and $j\sim \{1,\ldots,m\}$ chosen uniformly at random. Let $\ell$ be the line through $u$ in the $j$-th direction. Let $e_j=(0,\ldots,0,1,0,\ldots,0)\in \F_q^m$, where the unique $1$ is in the $j$-th position. Choose a uniformly random $v\in \ell$ and $\gamma\in \F_q$ and check that 
		\[\sum_{i=0}^d g(u+ie_j,\kappa(\gamma \alpha_{u,v,i})) \,=\, g(v,\kappa(\gamma))\;,\]
		where $\alpha_{u,v,i}$ is defined in~\eqref{eq:interp-coeff}.
	\item \textbf{Hadamard test:} Let $u\sim\F_q^m$ be chosen uniformly at random and $\alpha,\beta\in \F_2^t$ chosen uniformly at random. Check that 
	\[g(u,\alpha)+g(u,\beta)\,=\,g(u,\alpha+\beta)\;.\] 	
    \end{enumerate}
  \end{gamespec}
  \caption{A local test for $\code_{\bRM}$}
  \label{fig:bRM-tester}
\end{figure}

We do not know if $G_\bRM = \Z_2^K$, with $K=t(d+1)^m$. Instead we modify the presentation $G(h_\bRM)$ by adding pairwise commutation relations in a similar manner as Lemma~\ref{lem:com-code}. Let 
\[ G(h_\bRM) \,=\, \big\langle x_1,\ldots,x_N \;:\; \{R^\sq_k\}\,,\; \{R^\ld_k\}\,,\; \{R^\had_{k}\} \big\rangle\;,\]
where  $R^\sq_k$ ranges overal all relations of the form $x_i^2=e$ for $i\in \{1,\ldots,N\}$, $R^\ld_k$ ranges over all relations implied by the ``low-degree test'' in Figure~\ref{fig:bRM-tester} and $R^\had_{k}$ ranges over all the relations implied by the ``Hadamard test.'' For $k=(i,j)\in\{1,\ldots,N\}$ such that $i<j$ let $R^\com_k$ be the relation $[x_i,x_j]=e$.
Then we define
\begin{equation}\label{eq:z2k-eff}
 G' \,=\,\big\langle x_1,\ldots,x_N \;:\; \{R^\sq_k\}\,,\; \{R^\ld_k\}\,,\; \{R^\had_{k}\}\, , \; \{ R^\com_k\}\big\rangle\;.
\end{equation}
From Lemma~\ref{lem:com-code} it follows that $G'=\Z_2^K$. Our main result is an efficient stability result for this presentation. To state this we need to introduce distributions $\mu_S$ and $\mu_R$ on the generators and relations of $G'$. The distribution $\mu_S$ is taken to be the uniform distribution over $[N] = (\F_2^t)^m \times \F_2^t$. The distribution $\mu_R$ is obtained as follows. With probability $1/4$ each, a relation from $\{R^\sq_k\}$, $\{R^\ld_k\}$ or $\{R^\had_k\}$ is chosen uniformly at random. With probability $1/4$, a random commutation relation from $R^\com_k$ is chosen according to the uniform mixture of the following two distributions:
\begin{enumerate}
\item For the first distribution, we select $u\in \F_q^m$ uniformly at random, $j\in\{1,\ldots,m\}$ uniformly at random, and $i\neq i'\in\{0,\ldots,d\}$ uniformly at random. Finally, select $\alpha,\beta\in \F_2^t$ uniformly at random and check commutation between $x_{u+i e_j,\alpha}$ and $x_{u+i' e_j,\beta}$. 
\item For the second distribution, we first select $j\in\{1,\ldots,m\}$ and $u_{m-j+2},\ldots,u_m \in \F_2^t$ uniformly at random. Then select $v,v' \in (\F_2^t)^m$ uniformly at random, conditioned on the last $(j-1)$ coordinates matching $u_{m-j+2},\ldots,u_m$. Finally, select $\alpha,\beta\in \F_2^t$ uniformly at random and check commutation between $x_{u,\alpha}$ and $x_{u,\beta}$. 
\end{enumerate}

@@

\begin{proof}
Let $(\mM,\tau)$ be a tracial von Neumann algebra and $\phi$ be an $\eps$-homomorphism of $\langle S:R\rangle$ on $(\mU(\mM),\tau)$. 
 Here, $S = \{s_{u,\alpha}: u\in (\F_2^t)^m, \alpha\in \F_2^t\}$ and $R$ is the set of all relations in~\eqref{eq:z2k-eff}, i.e.\ $R= \{R_k^\sq\}\cup \{R^\ld_k\}\cup\{R^\had_k\}\cup\{R^\com_k\}$. Let $N=|S|$.

We first exploit the relations $R_k^\sq$ to show that we may assume without loss of generality that $\phi$ sends each element of $S$ to a Hermitian involution.

\begin{claim}\label{claim:z2-stab-1}
There is an $\eps^{(1)}=O(d\eps)$ and an $\eps^{(1)}$-homomorphism $\phi^{(1)}$ of $\langle S:R\rangle$ on $(\mU(\mM),\tau)$ such that $\phi^{(1)}(x)$ is a Hermitian involution for all $x\in S$, and furthermore
\begin{equation}\label{eq:z2-stab-1a}
 \Es{x\sim\mu_S} \big\| \phi(x) - \phi^{(1)}(x) \big\|_\tau^2 \,\leq\, \eps^{(1)}\;.
\end{equation}
\end{claim}

\begin{proof}
Using elementary calculations (see e.g.~\cite[Lemma 3.6]{slofstra2019set}) we see that for any complex $\alpha$, 
\[ \big| \sgn\Re\alpha-\alpha\big| \,\leq\, \Big(1+\frac{1}{\sqrt{2}}\Big) \big|\alpha^2 -1 \big|\;.\]
For any $i\in \{1,\ldots,N\}$ let $\phi^{(1)}(x_i) = \sgn\Re (\phi(x_i))$. Then the claim follows since 
\[ \Es{i\in\{1,\ldots,N\}} \big\| \phi(x_i)^2-\Id \big\|_\tau^2 \,\leq\, 4\eps\;,\]
by assumption and the definition of $\mu_R$, which places weight $\frac{1}{4}$ on the relations in $R_k^\sq$. Since $\mu_S$ is uniform on $S$,~\eqref{eq:z2-stab-1a} follows. Furthermore, because all relations in $R$ have length at most $O(d)$, $\phi^{(1)}$ is an $O(d\eps)$-homomorphism of $\langle S:R\rangle$\footnote{This requires more explanation, e.g.\ we need the marginal of any variable appearing in a relation to be uniform. I'm also not sure it's not $d^2$ instead of $d$.}. 
\end{proof}

For ease of notation we relabel $\phi^{(1)}$ and $\eps^{(1)}$ as $\phi$ and $\eps$ respectively. Next we exploit the relations $\{R^\had_k\}$ to show the following.

\begin{claim}\label{claim:z2-stab-2}
Let $q=2^t$. 
For every $u\in \F_q^m$ there is a projective measurement $\{P^u_\alpha\}_{\alpha\in \F_{q}}$ on $\mM$ such that 
\begin{equation}\label{eq:z2-stab-2}
 \Es{u\in \F_q^m} \Es{\alpha\in \F_2^t} \Big\| \phi(s_{u,\alpha}) - \sum_{\beta\in\F_q} (-1)^{\alpha \cdot \kappa(\beta)} P^u_\beta \Big\|_\tau^2 \,=\, O(\eps)\;. 
\end{equation}
\end{claim}

\begin{proof}
Since $\mu_R$ places weight $1/4$ on relations $\{R^\had_k\}$ we deduce that 
\begin{equation}\label{eq:stab-rm-1}
\Es{u\in \F_q^m} \Es{\alpha,\beta\in \F_2^t} \big\|\phi(s_{u,\alpha})\phi(s_{u,\beta})\phi(s_{u,\alpha+\beta})-\Id\big\|_\tau^2 \,\leq\, 4\eps\;. 
\end{equation}
Fix an $u\in \F_q^m$ and apply Lemma~\ref{lem:had-stab} for that $u$. This gives a partial isometry $w\in P\mM_\infty\Id_\mM$ and a representation $\{U_\alpha\}$ of $\F_2^t$ on $\mN=P\mM_\infty P$ such that 
\begin{equation}\label{eq:stab-rm-1b}
 \Es{\alpha \in \F_2^t} \big\| \phi(s_{u,\alpha}) - w^* U_\alpha w \big\|_\tau^2 \,=\, O(\eps_u)\;,
\end{equation}
where $\eps_u\geq 0$ is such that $\Es{u\in \F_q^m} \eps_u = 4\eps$. Because $\{U_\alpha\}$ are a representation of the abelian group $\F_2^t$, there is a projective measurement $\{Q^u_\beta\}_{\beta\in \F_2^t}$ on $\mN$ such that $U_\alpha = \sum_\beta (-1)^{\alpha\cdot \beta} Q^u_\beta$. ($\{Q^u_\beta\}$ can be found explicitly by applying the Fourier transform.) Applying Lemma~\ref{lem:pull-back}, we deduce a projective measurement $\{P^u_\beta\}_{\beta\in \F_q}$  on $\mM$ such that (by the triangle inequality)
\begin{align*}
 \Es{\alpha \in \F_2^t} \Big\| \phi(s_{u,\alpha}) - \sum_{\beta\in\F_q} (-1)^{\alpha \cdot \kappa(\beta)} P^u_\beta \Big\|_\tau^2
&\leq 2 \Es{\alpha \in \F_2^t} \Big( \Big\| \phi(s_{u,\alpha}) - w^*\Big(\sum_{\beta\in\F_2^t} (-1)^{\alpha \cdot (\beta)} Q^u_\beta \Big) w \Big\|_\tau^2 \\
&\qquad\qquad+   \Big\|  \sum_{\beta\in\F_2^t} (-1)^{\alpha \cdot \beta} w^* Q^u_\beta w - \sum_{\beta\in\F_q} (-1)^{\alpha \cdot \kappa(\beta)} P^u_\beta \Big\|_\tau^2\Big)\\
 &\leq\, O(\eps_u) +  \sum_{\beta\in\F_q} \Big\| w^* Q^u_{\kappa(\beta)} w - P^u_{\beta} \Big\|_\tau^2\\
&\leq O(\eps_u)\;,
\end{align*}
where the first inequality follows from the triangle inequality, the second is by~\eqref{eq:stab-rm-1b} for the first term and Parseval's identity for the second, and the last by the guarantees obtained from Lemma~\ref{lem:pull-back}.
\end{proof}

For $u\in \F_q^m$ let $\{P^{u}_\beta\}_{\beta \in \F_q}$ be the projective measurement obtained from Claim~\ref{claim:z2-stab-2}, and 
\[ U_u = \sum_{\beta\in\F_q} \omega_q^\beta P^u_{\beta}\;,\]
where $\omega_q = e^{2i\pi/q}$. Then $U_u \in \mU(\mM)$.

%We observe that
%\begin{equation}\label{eq:z2-stab-2a}
%\Es{u\in\F_q^m}\Es{\alpha\in \F_q} \big\| U_u^\alpha - \phi(s_{u,\kappa(\alpha)})\big\|_\tau^2 \,=\, O(\eps)\;.
%\end{equation}
%To see this, observe that for any $\alpha\in \F_q$, $\omega_q^\alpha = (-1)^{\tr(\alpha)}$. Thus from the definition of $U_u$ we get
%\begin{align*}
%U_u^\alpha &= \sum_\beta \omega_q^{\alpha\beta} P^u_\alpha\\
 %&= \sum_\beta (-1)^{\tr(\alpha\beta)} P^u_\alpha\\
%&= \sum_\beta (-1)^{\alpha\cdot \beta} P^u_\alpha\;,
%\end{align*}
%where in the last line we slightly abused notation and identified $\alpha,\beta\in\F_q$ with $\kappa(\alpha),\kappa(\beta)\in \F_2^t$ respectively. \eqref{eq:z2-stab-2a} then follows from~\eqref{eq:z2-stab-2}.

\begin{claim}\label{claim:z2-stab-2b}
For $u\in \F_q^m$, $j\in\{1,\ldots,m\}$ and $i\in \{0,\ldots,d\}$ let $U_{i} = U_{u+ie_j}$ and, for all $\beta\in\F_q$, $P^{(i)}_\beta = P^{u+ie_j}_\beta$ (leaving the dependence on $u$ implicit). Then 
\begin{equation}\label{eq:z2-stab-2b-0a}
\Es{u\in \F_q^m} \Es{\substack{j\in\{1,\ldots,m\}\\i\neq i' \in \{0,\ldots,d\}}}\Es{\alpha,\alpha'\in \F_q} \big\| \big[ U_i^\alpha, U_{i'}^{\alpha'}\big]-\Id\big\|_\tau^2\,=\, O\big(\sqrt{\eps}\big)\;, 
\end{equation}
and
\begin{equation}\label{eq:z2-stab-2b-0b}
 \Es{u\in \F_q^m} \Es{j\in\{1,\ldots,m\}} \Es{\substack{v\in \F_q^m \\v_{m-j+2}=u_{m-j+2},\ldots,v_m=u_m}}\Es{\alpha,\alpha'\in \F_q} \big\| \big[ U_u^\alpha, U_{v}^{\alpha'}\big]-\Id\big\|_\tau^2\,=\, O\big(\sqrt{\eps}\big)\;,
\end{equation}
where the expectation is over a uniformly random $u$ and $j$, and a uniformly random $v$ conditioned on its last $(j-1)$ coordinates matching those of $u$. 
\end{claim}

\begin{proof}
Due to the test of the relations $\{R^\com_k\}$, it holds that \tnote{this will depend on the distribution placed on commutation relations; the factor $4$ might be off} 
\begin{align}
\Es{u\in \F_q^m} \Es{\substack{j\in\{1,\ldots,m\} \\ i\neq i' \in \{0,\ldots,d\}}} \Es{\alpha,\beta\in \F_q} \big\| [\phi(s_{u+i e_j,\alpha}),\phi(s_{u+i' e_j,\beta})] - \Id \big\|_\tau^2&\leq 8\eps\; ,\label{eq:z2-stab-2b-1}\\
\Es{u\in \F_q^m} \Es{j\in\{1,\ldots,m\}} \Es{\substack{v\in \F_q^m \\v_{m-j+2}=u_{m-j+2},\ldots,v_m=u_m}}\Es{\alpha,\alpha'\in \F_q} \big\| [\phi(s_{u,\alpha}),\phi(s_{v,\beta})]-\Id\big\|_\tau^2&\leq 8\eps\;.\label{eq:z2-stab-2b-1b}
\end{align}
Now, for any $u,v\in \F_q^m$
\begin{align*}
\Es{\alpha,\alpha'\in \F_q} &\tau\big( U_u^\alpha U_{v}^{\alpha'} U_u^{-\alpha} U_{v}^{-\alpha'} \big) \\
&= \sum_{\beta,\beta',\gamma,\gamma'\in \F_q}\Es{\alpha\in \F_q} \omega_q^{\alpha(\beta-\beta')} \Es{\alpha'\in \F_q} \omega_q^{\alpha'(\gamma-\gamma')}  \tau(P^{u}_\beta P^{v}_\gamma P^{u}_{\beta'} P^{v}_{\gamma'}\big)\\
&=\sum_{\beta,\beta',\gamma,\gamma'\in \F_q}\Es{\alpha\in\F_2^t} (-1)^{\alpha\cdot\kappa(\beta-\beta')} \Es{\alpha'\in\F_2^t} (-1)^{\alpha'\cdot\kappa(\gamma-\gamma')} \tau(P^{u}_\beta P^{v}_\gamma P^{u}_{\beta'} P^{v}_{\gamma'}\big)\\
&=\Es{\alpha,\alpha'\in\F_2^t}  \tau\Big( \Big(\sum_{\beta\in \F_q} (-1)^{\alpha\cdot \kappa(\beta)} P^{u}_\beta\Big) \Big(\sum_{\gamma\in \F_q} (-1)^{\alpha'\cdot \kappa(\gamma)} P^{v}_\gamma\Big)\Big(\sum_{\beta'\in \F_q} (-1)^{\alpha\cdot \kappa(\beta')} P^{u}_{\beta'}\Big)\Big(\sum_{\gamma'\in \F_q} (-1)^{\alpha'\cdot \kappa(\gamma')} P^{v}_{\gamma'}\Big)\Big)\;.
\end{align*}
Here, the second line follows since for $\beta,\beta'\in \F_q$,
\[ \Es{\alpha\in \F_q} \omega_q^{\alpha(\beta-\beta')}\,=\,\Es{\alpha\in\F_2^t} (-1)^{\alpha\cdot \kappa(\beta-\beta')}\,=\delta_{\beta,\beta'}\;.\]
It follows that, for any distribution on $(u,v)$ such that both marginals are uniform over $\F_q^m$,
\begin{align*}
\Es{u,v} \Es{\alpha,\alpha'\in \F_q} \big\| [U_u^\alpha ,U_{v}^{\alpha'} ]-\Id\big\|_\tau^2
&= \Es{\alpha,\alpha'\in \F_2^t} \Big\| \Big[\sum_{\beta\in \F_q} (-1)^{\alpha\cdot \kappa(\beta)} P^{u}_\beta ,\sum_{\beta'\in \F_2^t} (-1)^{\alpha\cdot \kappa(\beta')} P^{v}_{\beta'} \Big]-\Id\Big\|_\tau^2\\
&= \Es{\alpha,\alpha'\in \F_2^t} \big\| \big[ \phi(s_{u,\alpha}),\phi(s_{v,\alpha'})\big]-\Id\big\|_\tau^2+O\big(\sqrt{\eps}\big)\;,
\end{align*}
where the second line follows from Claim~\ref{claim:z2-stab-2}. Thus~\eqref{eq:z2-stab-2b-0a} follows from~\eqref{eq:z2-stab-2b-1}, and~\eqref{eq:z2-stab-2b-0b} follows from~\eqref{eq:z2-stab-2b-1b}.
\end{proof}

Now we show the following, which essentially follows from the previous claim.

\begin{claim}\label{claim:z2-stab-3}
For every $u\in \F_q^m$ and $j\in\{1,\ldots,m\}$, there is a projective measurement $\{R^{u,j}_\alpha\}_{\alpha\in\F_q^{d+1}}$ on $\mM$ such that 
\[ \Es{u\in \F_q^m} \Es{j\in\{1,\ldots,m\}} \Es{x_0,\ldots,x_{d}\in \F_q} \Big\| U_0^{x_0}\cdots U_d^{x_d} -  \sum_{\alpha} \omega^{\sum x_i\alpha_i} R^{u,j}_\alpha\Big\|_\tau^2 \,=\, \poly(d,\eps)\;.\]
Similarly, for any $u,v\in \F_q^m$ there is a projective measurement $\{R^{u,v}_{\alpha,\beta}\}_{(\alpha,\beta)\in\F_q^{2}}$ on $\mM$ such that 
\[  \Es{u\in \F_q^m} \Es{j\in\{1,\ldots,m\}} \Es{\substack{v\in \F_q^m \\v_{m-j+2}=u_{m-j+2},\ldots,v_m=u_m}} \Es{x,y\in \F_q} \Big\| U_u^{x}  U_v^{y} -  \sum_{\alpha} \omega^{ \alpha x + \beta y} R^{u,v}_{\alpha,\beta}\Big\|_\tau^2 \,=\, \poly(\eps)\;.\]
\end{claim}

\begin{proof}
We show the first part only, as the second part is entirely analogous. Fix an $u\in \F_q^m$ and a direction $j\in \{1,\ldots,m\}$. Define
$\psi: \Z_q^{d+1} \to \mU(\mM)$ by 
\[\psi(x_0,\ldots,x_{d}) \,=\, U_0^{x_0} \cdots U_{d}^{x_{d}}\;,\]
where for each $i\in\{0,\ldots,d\}$, $U_i$ is the order-$q$ unitary defined in Claim~\ref{claim:z2-stab-2b}. Using Claim~\ref{claim:z2-stab-2b} and the triangle inequality we verify that the map $\psi$ is a $\delta=O(d^4\sqrt{\eps})$-approximate homomorphism of $\Z_q^{d+1}$ on $\mU(\mA)$, with respect to the exhaustive presentation. To verify this first note that for any unitaries $V,W$, and any $i,i'$ we have that on average over $u$ and $j$,
\begin{align*}
\Es{\alpha,\alpha'\in \F_q} \big\| V U_i^\alpha U_{i'}^{\alpha'} W^* - V U_{i'}^{\alpha'} U_i^\alpha  W^*\big\|_\tau^2 \,=\, O\big(d^2 \sqrt{\eps}\big)\;,
\end{align*}
by Claim~\ref{claim:z2-stab-2b}, where the factor $d^2$ is because we require the relation to hold for all $i,i'$. Applying this relation $O(d^2)$ times gives
\begin{align*}
 \Es{x_0,\ldots,x_d\in\F_q}\Es{y_0,\ldots,y_d\in\F_q}\big\|\psi(x_0,\ldots,x_d)\psi(y_0,\ldots,y_d) -\psi(x_0+y_0,\ldots,x_d+y_d)\big\|_\tau^2 \,=\, O\big(d^4 \sqrt{\eps}\big)\;,
\end{align*}
as desired.
Thus we may apply Theorem~\ref{thm:gh} to obtain a partial isometry $w\in P\mM_\infty\Id_\mM$ and a family of commuting unitaries $\{V_i\}$ of order $q$ each on $\mN=P\mM_\infty P$ such that, on average over $u$ and $j$,
\begin{equation}\label{eq:z2-stab-3-1}
\Es{x_0,\ldots,x_{d}\in\F_q} \big\| \psi(x_0,\ldots,x_{d+1}) - w^* V_0^{x_0}\cdots V_{d}^{x_{d}} w \big\|_\tau^2 \,=\, \poly(d,\eps)\;.
\end{equation}
Because $\{V_i\}$ are a representation of $\Z_q^{d+1}$ (which is Abelian), there is a projective measurement $\{Q_\alpha\}_{\alpha\in \F_q^{d+1}}$ on $\mN$ such that for each $i\in\{0,\ldots,d\}$, 
\[ V_i \,=\, \sum_{\alpha\in \Z_q^{d+1}} \omega^{\alpha_i} Q_\alpha\;,\]
and  
\[ V_0^{x_0}\cdots V_{d}^{x_{d}}  \,=\, \sum_{\alpha\in \Z_q^{d+1}} \omega^{\sum x_i\alpha_i} Q_\alpha\;.\]
%($\{Q^u_\alpha\}$ can be found explicitly by applying the Fourier transform.) 
Applying Lemma~\ref{lem:pull-back}, we deduce a projective measurement $\{R^{u,j}_\alpha\}_{\alpha\in \Z_q^{d+1}}$  on $\mM$ such that by~\eqref{eq:z2-stab-3-1}, on average over $u$ and $j$,
\[  \Es{x_0,\ldots,x_{d}\in\F_q} \Big\| \psi(x_0,\ldots,x_{d+1}) -  \sum_{\alpha\in \Z_q^{d+1}} \omega^{\sum x_i\alpha_i} R_\alpha^{u,j}\Big\|_\tau^2 \,=\, \poly(d,\eps)\;.\]
\end{proof}

@@

\begin{claim}\label{claim:z2-stab-5}
Use the same notation as in Claim~\ref{claim:z2-stab-2b}. Let $u\in \F_q^m$, $\ell$ an axis-parallel line through $u$ and $v\in\F_q^m\cap \ell$. Let $(\alpha_{u,v,i})_{i=0,\ldots,d}$ be the interpolation coefficients~\eqref{eq:interp-coeff}. Let $U_v = \sum_\beta \omega^\beta P^v_\beta$. Then 
\begin{equation}\label{eq:z2-stab-5-0}
\Es{u\in\F_q^m} \Es{\ell: u\in \ell} \Es{v\in \ell} \Es{\gamma\in\F_q} \Big\| U_0^{\gamma \alpha_{u,v,0}}\cdots U_d^{\gamma  \alpha_{u,v,d}} - U_v^\gamma \big\|_\tau^2 \,=\, 1-O\big(d\sqrt{\eps}\big)\;,
\end{equation}
where the expectation is over a uniformly random axis-parallel line that contains $u$, and a uniformly random $v\in\ell$. 
\end{claim}
